Inferring Continuous and Discrete Population Genetic Structure Across Space

### 2017-08-18 writeup/conStruct.tex
intro changes

\newpage
\section*{Introduction}
The fundamental goal of evolutionary genetics is to link patterns of genetic variation 
to the processes that have generated and maintained them.
Often, the first step in the analysis of genetic data is to define evolutionary units of analysis
by delineating discrete populations from which individuals or alleles are sampled \citep{wright1949genetical}.
These units are useful for defining a scope of study in research on selection or local adaptation (ref),
or for identifying either locally adapted alleles (human skin color, ralph \& coop, gartersnakes) or alleles involved in disease (cystic fibrosis).
Delineating populations can also be useful for informing conservation priorities \citep{Moritz1994,Waples_1998,Moritz_etal_2002}.
However, if underlying population structure is continuous, this may entail a loss of power.

There have been many methods proposed to characterize populations,
including generating population phylogenies \citep{CavalliSforza1975, treemix},
$k$-means clustering analyses performed on dimensionality-reduction approaches (ref), 
such as principal components analysis \citep{menozzi1978synthetic,novembre_interpreting_2008, price2006eigenstrat},
and model-based clustering approaches \citep[e.g.][]{STRUCTURE, falush2003, hubisz2009,ADMIXTURE, FINESTRUCTURE, fastStructure, huelsenbeck2007inference, Corander2003,TESS,geneland}, among others.
Here, we will focus on model-based clustering, the most widely used class of approaches for population delineation.
Like the method we present here, these are based on a statistical model,
but do not perform inference directly with a demographic model.
Each of these methods perform best in particular situations,
but all can give misleading results when applied to data that show a continuous pattern of isolation by distance 
\citep{Wright1943, novembre_interpreting_2008, Frantz2009}.

>>>>

\newpage
\section*{Introduction}
A fundemental quandary in the description of biological diversity, is
the fact that diversity shows both discrete and continuous patterns. 
For example, there is disagreement on species concepts because the
process of speciation is usually gradual, and so there is no set point
in the continuous divergence of populations when they become discrete species.
This issue extends below the species level as patterns of phenotypic
and genetic diversity within and among populations are shaped
by continuous migration and drift, as well as more discrete events and
processes such as geographic barriers, rapid expansions, bottlenecks, and rare
long distance migration.

From a practical standpoint, even while acknowledging continuous
processes, we often need to identify somewhat separable populations 
from which individuals are sampled \citep{wright1949genetical}.
Delineating populations is useful for systematics and for
informing conservation priorities \citep{Moritz1994,Waples_1998,Moritz_etal_2002}.
Furthermore, we often need to identify sets of indivduals resulting from reasonably coherent
evolutionary histories for downstream analyses to learn about population history and adaptation.

%In studying populations we need group individuals 
%Many downstream

%Biology diversity

%The fundamental goal of evolutionary genetics is to link patterns of genetic variation 
%to the processes that have generated and maintained them.
%Often, the first step in the analysis of genetic data is to define evolutionary units of analysis
%by delineating discrete populations from which individuals or alleles are sampled \citep{wright1949genetical}.
%These units are useful for defining a scope of study in research on selection or local adaptation (ref),
%or for identifying either locally adapted alleles (human skin color, ralph \& coop, gartersnakes) or alleles involved in disease (cystic fibrosis).
%Delineating populations can also be useful for informing conservation priorities \citep{Moritz1994,Waples_1998,Moritz_etal_2002}.
%However, if underlying population structure is continuous, this may entail a loss of power.

There have been many methods proposed to characterize population
genetic structure,
including generating population phylogenies \citep{CavalliSforza1975, treemix},
$k$-means clustering analyses performed on dimensionality-reduction approaches (ref), 
such as principal components analysis \citep{menozzi1978synthetic,novembre_interpreting_2008, price2006eigenstrat},
and model-based clustering approaches \citep[e.g.][]{STRUCTURE, falush2003, hubisz2009,ADMIXTURE, FINESTRUCTURE, fastStructure, huelsenbeck2007inference, Corander2003,TESS,geneland}, among others.
Here, we will focus on model-based clustering, the most widely used class of approaches for population delineation.
Like the method we present here, these are based on a statistical model,
but do not perform inference directly with a demographic model.
Each of these methods perform best in particular situations,
but all can give misleading results when applied to data that show a continuous pattern of isolation by distance 
\citep{Wright1943, novembre_interpreting_2008, Frantz2009}.
